---
title: "Data prep"
author: "BVD"
date: "6/23/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load data
```{r message=FALSE}
load("~/DFW7_1_19.RData")
DFW<- DFW7_1_19
```

Load packages
```{r message=FALSE}
library("cAIC4")
library("MuMIn")
library("lme4")
library("dummies")
library(foreach)
library(ggplot2)
library(multcomp)

library(lattice)
library(hglm)
library(gmodels)
library(frequency)

library("dplyr")
plot_col <- c('#66c2a5', '#fc8d62', '#8da0cb')
#cbbpalette <- c('#000000','#E69F00','#56B4E9')
cbbpalette <- c( "#009E73", "#e79f00", "#9ad0f3", "#0072B2", "#D55E00", 
    "#CC79A7","#000000", "#F0E442") #colorblind and grayscale friendly.
```

```{r}

```

```{r}
logit2prob <- function(logit){
  odds <- exp(logit)
  prob <- odds / (1 + odds)
  return(prob)
}

logit2prob(-0.75) #(White men without LAs)
logit2prob(-0.5) #(White men with LAs)

model_probs <- function(courses){
probcourses <- summary(courses)

temp <- data.frame(group=rownames(probcourses$linfct),
             q = probcourses$test$coefficients, 
             se = (probcourses$test$sigma))

temp$UL <- temp$q+temp$se
temp$LL <- temp$q-temp$se
temp2 <- logit2prob(temp[c(2,4,5)])
return <- temp2
}
```

ID LA courses
```{r}
DFW$LA<-0
DFW$LA[
  DFW$INSTRUCTOR1_NAME=="Brookes,David T" & DFW$Number=="202A" & DFW$TERM=="2172" | 
         DFW$INSTRUCTOR1_NAME=="Brookes,David T" & DFW$Number=="202A" & DFW$TERM=="2178" | 
         DFW$INSTRUCTOR1_NAME=="Brookes,David T" & DFW$Number=="202A" & DFW$TERM=="2182" | 
         DFW$INSTRUCTOR1_NAME=="Brookes,David T" & DFW$Number=="202A" & DFW$TERM=="2188" | 
         DFW$INSTRUCTOR1_NAME=="Brookes,David T" & DFW$Number=="202A" & DFW$TERM=="2192" | 
         DFW$INSTRUCTOR1_NAME=="Brookes,David T" & DFW$Number=="202B" & DFW$TERM=="2168" |
         DFW$INSTRUCTOR1_NAME=="Arpin,Paul C" & DFW$Number=="204B" & DFW$TERM=="2172" | 
         DFW$INSTRUCTOR1_NAME=="Arpin,Paul C" & DFW$Number=="204B" & DFW$TERM=="2178" | 
         DFW$INSTRUCTOR1_NAME=="Arpin,Paul C" & DFW$Number=="204B" & DFW$TERM=="2192" |
    DFW$INSTRUCTOR1_NAME=="Arpin,Paul C" & DFW$Number=="204A" & DFW$TERM=="2188" |
    DFW$INSTRUCTOR1_NAME=="Arpin,Paul C" & DFW$Number=="202B" & DFW$TERM=="2182" |
    DFW$INSTRUCTOR1_NAME=="Ayars,Eric J" & DFW$Number=="204A" & DFW$TERM=="2172" |
    DFW$INSTRUCTOR1_NAME=="Ayars,Eric J" & DFW$Number=="204A" & DFW$TERM=="2178" |
    DFW$INSTRUCTOR1_NAME=="Ayars,Eric J" & DFW$Number=="204A" & DFW$TERM=="2182" |
    DFW$INSTRUCTOR1_NAME=="Bailey,Paul L" & DFW$Number=="118" & DFW$TERM=="2178" |
    DFW$INSTRUCTOR1_NAME=="Bailey,Paul L" & DFW$Number=="118" & DFW$TERM=="2182" |
    DFW$INSTRUCTOR1_NAME=="Bailey,Paul L" & DFW$Number=="119" & DFW$TERM=="2168" |
    DFW$INSTRUCTOR1_NAME=="Bailey,Paul L" & DFW$Number=="119" & DFW$TERM=="2172" |
    DFW$INSTRUCTOR1_NAME=="Bailey,Paul L" & DFW$Number=="119" & DFW$TERM=="2178" |
    DFW$INSTRUCTOR1_NAME=="Bailey,Paul L" & DFW$Number=="119" & DFW$TERM=="2182" |
    DFW$INSTRUCTOR1_NAME=="Bailey,Paul L" & DFW$Number=="119" & DFW$TERM=="2188" |
    DFW$INSTRUCTOR1_NAME=="Bailey,Paul L" & DFW$Number=="120" & DFW$TERM=="2192" |
    DFW$INSTRUCTOR1_NAME=="Danner,Richard S" & DFW$Number=="118" & DFW$TERM=="2182" |
    DFW$INSTRUCTOR1_NAME=="Danner,Richard S" & DFW$Number=="120" & DFW$TERM=="2172" |
    DFW$INSTRUCTOR1_NAME=="Danner,Richard S" & DFW$Number=="118" & DFW$TERM=="2192" |
    DFW$INSTRUCTOR1_NAME=="Herrera,Christine A" & DFW$Number=="120" & DFW$TERM=="2172" |
    DFW$INSTRUCTOR1_NAME=="Herrera,Christine A" & DFW$Number=="120" & DFW$TERM=="2182" |
    DFW$INSTRUCTOR1_NAME=="Herrera,Christine A" & DFW$Number=="120" & DFW$TERM=="2188" |
    DFW$INSTRUCTOR1_NAME=="Herrera,Christine A" & DFW$Number=="120" & DFW$TERM=="2192" |
    DFW$INSTRUCTOR1_NAME=="Lin,Yuhfen" & DFW$Number=="202A" & DFW$TERM=="2168" |
    DFW$INSTRUCTOR1_NAME=="Lin,Yuhfen" & DFW$Number=="202A" & DFW$TERM=="2188" |
    DFW$INSTRUCTOR1_NAME=="Lin,Yuhfen" & DFW$Number=="202A" & DFW$TERM=="2192" |
    DFW$INSTRUCTOR1_NAME=="Lin,Yuhfen" & DFW$Number=="202B" & DFW$TERM=="2172" |
    DFW$INSTRUCTOR1_NAME=="Lin,Yuhfen" & DFW$Number=="202B" & DFW$TERM=="2178" |
    DFW$INSTRUCTOR1_NAME=="Lin,Yuhfen" & DFW$Number=="202B" & DFW$TERM=="2182" |
    DFW$INSTRUCTOR1_NAME=="Lin,Yuhfen" & DFW$Number=="204B" & DFW$TERM=="2178" |
    DFW$INSTRUCTOR1_NAME=="Newton,Mark H" & DFW$Number=="343" & DFW$TERM=="2188" |
    DFW$INSTRUCTOR1_NAME=="Newton,Mark H" & DFW$Number=="343" & DFW$TERM=="2192" |
    DFW$INSTRUCTOR1_NAME=="Pechkis,Hyewon K" & DFW$Number=="204B" & DFW$TERM=="2182" |
    DFW$INSTRUCTOR1_NAME=="Pechkis,Hyewon K" & DFW$Number=="204B" & DFW$TERM=="2188" |
    DFW$INSTRUCTOR1_NAME=="Pechkis,Hyewon K" & DFW$Number=="204B" & DFW$TERM=="2192" |
    DFW$INSTRUCTOR1_NAME=="Petrova-Mayor,Anna I" & DFW$Number=="204A" & DFW$TERM=="2178" |
    DFW$INSTRUCTOR1_NAME=="Petrova-Mayor,Anna I" & DFW$Number=="204C" & DFW$TERM=="2172" |
    DFW$INSTRUCTOR1_NAME=="Petrova-Mayor,Anna I" & DFW$Number=="204C" & DFW$TERM=="2178" |
    DFW$INSTRUCTOR1_NAME=="Petrova-Mayor,Anna I" & DFW$Number=="204C" & DFW$TERM=="2182" |
    DFW$INSTRUCTOR1_NAME=="Strand,Krista L" & DFW$Number=="110" & DFW$TERM=="2182" |
    DFW$INSTRUCTOR1_NAME=="Strand,Krista L" & DFW$Number=="110" & DFW$TERM=="2188" |
    DFW$INSTRUCTOR1_NAME=="Tran,Cawa" & DFW$Number=="153" & DFW$TERM=="2188" |
    DFW$INSTRUCTOR1_NAME=="Tran,Cawa" & DFW$Number=="153" & DFW$TERM=="2192" |
    DFW$INSTRUCTOR1_NAME=="Van Dusen,Benjamin C" & DFW$Number=="141" & DFW$TERM=="2168" |
    DFW$INSTRUCTOR1_NAME=="Van Dusen,Benjamin C" & DFW$Number=="141" & DFW$TERM=="2172" |
    DFW$INSTRUCTOR1_NAME=="Van Dusen,Benjamin C" & DFW$Number=="141" & DFW$TERM=="2178" |
    DFW$INSTRUCTOR1_NAME=="Van Dusen,Benjamin C" & DFW$Number=="141" & DFW$TERM=="2182" |
    DFW$INSTRUCTOR1_NAME=="Van Dusen,Benjamin C" & DFW$Number=="141" & DFW$TERM=="2188" |
    DFW$INSTRUCTOR1_NAME=="Van Dusen,Benjamin C" & DFW$Number=="141" & DFW$TERM=="2192" |
    DFW$INSTRUCTOR1_NAME=="Van Dusen,Benjamin C" & DFW$Number=="321" & DFW$TERM=="2178" |
    DFW$INSTRUCTOR1_NAME=="Van Dusen,Benjamin C" & DFW$Number=="350" & DFW$TERM=="2182" |
    DFW$INSTRUCTOR1_NAME=="Zou,Xueli" & DFW$Number=="204A" & DFW$TERM=="2188" |
    DFW$INSTRUCTOR1_NAME=="Zou,Xueli" & DFW$Number=="204A" & DFW$TERM=="2192" |
    DFW$INSTRUCTOR1_NAME=="Pechkis,Joseph A" & DFW$Number=="202B" & DFW$TERM=="2192"
         ] <- 1

```

variables I care about
```{r}
DFW <- DFW[!is.na(DFW$Grade_Group),]
#DFW$DFW<-ifelse(DFW$Grade_Group=="A",0,ifelse(DFW$Grade_Group=="B",0,ifelse(DFW$Grade_Group=="C",0,ifelse(DFW$Grade_Group=="D",0,ifelse(DFW$Grade_Group=="CR",0,1)))))
DFW$DDFW<-ifelse(DFW$Grade_Group=="A",0,ifelse(DFW$Grade_Group=="B",0,ifelse(DFW$Grade_Group=="C",0,ifelse(DFW$Grade_Group=="CR",0,1))))
DFW$WD<-ifelse(DFW$Grade_Group=="A",0,
               ifelse(DFW$Grade_Group=="B",0,
                      ifelse(DFW$Grade_Group=="C",0,
                             ifelse(DFW$Grade_Group=="CR",0,ifelse(DFW$Grade_Group=="F",0,1)))))

names(DFW)[names(DFW)=="INSTRUCTOR1_NAME"] <- "instructor"
names(DFW)[names(DFW)=="FirstGeneration"] <- "FG_Status"
names(DFW)[names(DFW)=="COURSE_ID"] <- "course"
names(DFW)[names(DFW)=="Ethnicity"] <- "race"
names(DFW)[names(DFW)=="GenderLabel"] <- "gender"
DFW$courID <- paste(DFW$Subject,DFW$course,DFW$Section,DFW$TERM, sep='')

reduced <- as.data.frame(dplyr::select(DFW, TERM, Subject, Number,courID, Section, instructor,FG_Status, gender, race, dependent_income, LA, DDFW, WD))

```

creating dummy variables
```{r}
reduced <- dummy.data.frame(reduced, names=c("Number","race","FG_Status","gender"),sep = "_")
reduced <- cbind(reduced, dplyr::select(DFW, "Number","race","FG_Status","gender"))

reduced$nonwhite <- ifelse(reduced$race=="White",0,1)
reduced$algebra <- ifelse(reduced$Number=="202A" |DFW$Number=="202B",1,0)
reduced$Female <- ifelse(reduced$gender=="Female",1,0)
reduced$NULLGender <- ifelse(reduced$gender=="NULL",1,0)
reduced$FG_Status <- ifelse(reduced$FG_Status=="First Generation Student",1,0)
```

subset data
```{r}
red_phys <- reduced[reduced$Subject=="PHYS",]
red_math <- reduced[reduced$Subject=="MATH",]
red_math_phys <- reduced[reduced$Subject=="PHYS" | reduced$Subject=="MATH",]
red_202A <- reduced[reduced$Number=="202A",]
red_202B <- reduced[reduced$Number=="202B",]
red_204A <- reduced[reduced$Number=="204A",]
red_204B <- reduced[reduced$Number=="204B",]
red_204C <- reduced[reduced$Number=="204C",]
red_introPhys <- reduced[reduced$Number=="204A" | reduced$Number=="202A",]
red_202_204 <- reduced[reduced$Number=="204A" | reduced$Number=="204B" | reduced$Number=="204C" | reduced$Number=="202A" | reduced$Number=="202B",]
```

Simple model - intro phys
```{r}
intro_mod_simple <- (DDFW ~ LA + (1|courID))
intro_m_simple <- glmer(intro_mod_simple, data=red_introPhys,family = binomial)

summary(intro_m_simple)
ICC <- 0.3102/(0.3102+pi^2/3)
```

Find best model (intro physics) (https://quantscience.rbind.io/2017/12/26/model-selection-for-multilevel-modeling/)
```{r}
#m_all <- (DDFW ~ FG_Status*LA + Female*LA + nonwhite*LA + algebra*LA + (1|courID))
#m_all <- (DDFW ~ FG_Status*LA*Female*nonwhite + algebra + (1|courID))
m_all <- (DDFW ~ FG_Status*LA + Female*LA + nonwhite*LA + FG_Status*nonwhite + NULLGender*LA + (1|courID))
options(na.action = "na.fail")
intro_mod_dredge <- glmer(m_all, data=red_introPhys,family = binomial)
#intro_dredge <- dredge(intro_mod_dredge)
#model.sel(intro_dredge, rank=AICc)
```

Find best model (between classes) (https://quantscience.rbind.io/2017/12/26/model-selection-for-multilevel-modeling/)
```{r}
m_courses <- (DDFW ~ Number_204B*LA + Number_204C*LA + Number_202A*LA + Number_202B*LA + (1|courID))
options(na.action = "na.fail")
#phys_mod_dredge <- glmer(m_courses, data=red_202_204,family = binomial)
#phys_dredge <- dredge(phys_mod_dredge)
#model.sel(phys_dredge, rank=AIC)
```

running best model - intro phys
```{r}
intro_mod_uncond <- (DDFW ~ + (1|courID))
intro_m_uncond <- glmer(intro_mod_uncond, data=red_introPhys,family = binomial)

intro_mod_overall <- (DDFW ~ LA + (1|courID))
intro_m_overall <- glmer(intro_mod_overall, data=red_introPhys,family = binomial)

#intro_mod_finalold <- (DDFW ~ FG_Status*LA + Female*LA + nonwhite*LA + FG_Status*nonwhite + (1|courID))
#intro_m_finalold <- glmer(intro_mod_finalold, data=red_introPhys,family = binomial)
intro_mod_final <- (DDFW ~ FG_Status*LA + Female + nonwhite*LA + FG_Status*nonwhite + NULLGender + (1|courID))
intro_m_final <- glmer(intro_mod_final, data=red_introPhys,family = binomial)
summary(intro_m_uncond)
summary(intro_m_overall)
#summary(intro_m_finalold)
#summary(intro_m_final1)
summary(intro_m_final)
ICC <- 0.3102/(0.3102+pi^2/3)
```
Intro_m_final is our final model in line 218

running best model - intro phys (disaggregated)
These are not the final models
```{r}
intro_mod_uncond <- (DDFW ~ + (1|courID))
intro_m_uncond <- glmer(intro_mod_uncond, data=red_introPhys,family = binomial)

intro_mod_overall <- (DDFW ~ LA + (1|courID))
intro_m_overall <- glmer(intro_mod_overall, data=red_introPhys,family = binomial)

intro_mod_disag_final <- (DDFW ~ Number_202A*(FG_Status*LA + Female*LA + nonwhite*LA + FG_Status*nonwhite) + (1|courID))
intro_m_disag_final <- glmer(intro_mod_disag_final, data=red_introPhys,family = binomial)
summary(intro_m_disag_final)
summary(intro_m_final)
ICC <- 0.3102/(0.3102+pi^2/3)
```

running best model - across physics courses
```{r}
phys_mod_final <- (DDFW ~ Number_204B*LA + Number_204C*LA + Number_202A*LA + Number_202B*LA + (1|courID))
phys_m_final <- glmer(phys_mod_final, data=red_202_204,family = binomial)
summary(phys_m_final)
```

Logits
```{r}
#simple
#(Int, LA)
trad= c(1,0)
LAsup=c(1,1)

simple_example <- rbind (trad, LAsup)
simple2 <- glht(intro_m_simple, simple_example)

#courses
#(Int, 204B,LA,204C, 202A,202B,204B:LA ,204C:LA,202A:LA,202B:LA)
p204A=    c(1, 0,0,0, 0,0,0, 0,0,0)
p204ALA=  c(1, 0,1,0, 0,0,0, 0,0,0)
p204B=    c(1, 1,0,0, 0,0,0, 0,0,0)
p204BLA=  c(1, 1,1,0, 0,0,1, 0,0,0)
p204C=    c(1, 0,0,1, 0,0,0, 0,0,0)
p204CLA=  c(1, 0,1,1, 0,0,0, 1,0,0)
p202A=    c(1, 0,0,0, 1,0,0, 0,0,0)
p202ALA=  c(1, 0,1,0, 1,0,0, 0,1,0)
p202B=    c(1, 0,0,0, 0,1,0, 0,0,0)
p202BLA=  c(1, 0,1,0, 0,1,0, 0,0,1)

contrast_forms_course <- rbind ('p204A'=p204A,
                                'p204A LA'=p204ALA,
                                'p204B'=p204B,
                                'p204B LA'=p204BLA,
                                'p204C'=p204C,
                                'p204C LA'=p204CLA,                                
                                'p202A'=p202A,
                                'p202A LA'=p202ALA,
                                'p202B'=p202B,
                                'p202B LA'=p202BLA)

names(contrast_forms_course) <- c("Number_204B", "LA", "Number_204C", "Number_202A", "Number_202B", "Nmb_204B:LA", "LA:Nmb_204C", "LA:Nmb_202A", "LA:Nmb_202B")

courses <- glht(phys_m_final, contrast_forms_course)

#overall
#(Int, LA)
trad = c(1,0)
LA = c(1,1)

contrast_forms_overall <- rbind ('Traditional'=trad,
                                 'LA'=LA)
overall <- glht(intro_m_overall, contrast_forms_overall)

#Students
#(Int, fg,LA,fem,nwhi,Null, fg:LA,nwhi:LA, fg:nwhi)
WMR =   c(1, 0,0,0,0,0, 0,0, 0) #White, male, returning, no LA
WMRL =  c(1, 0,1,0,0,0, 0,0, 0) #White, male, returning, LA
BMR =   c(1, 0,0,0,1,0, 0,0, 0) 
BMRL =  c(1, 0,1,0,1,0, 0,1, 0)
WFR =   c(1, 0,0,1,0,0, 0,0, 0) 
WFRL =  c(1, 0,1,1,0,0, 1,0, 0)
WMF =   c(1, 1,0,0,0,0, 0,0, 0) 
WMFL =  c(1, 1,1,0,0,0, 1,0, 0)
WFF =   c(1, 1,0,1,0,0, 0,0, 0) 
WFFL =  c(1, 1,1,1,0,0, 1,0, 0)
BFR =   c(1, 0,0,1,1,0, 0,0, 0) 
BFRL =  c(1, 0,1,1,1,0, 0,1, 0)
BMF =   c(1, 1,0,0,1,0, 0,0, 1) 
BMFL =  c(1, 1,1,0,1,0, 1,1, 1)
BWF =   c(1, 1,0,1,1,0, 0,0, 1) 
BWFL =  c(1, 1,1,1,1,0, 1,1, 1)


contrast_forms_students <- rbind ('White male'=WMR,
                                'White male with LA'=WMRL,
                                'White female'=WFR,
                                'White female LA'=WFRL,                                
                                'Non-White male'=BMR,
                                'Non-White male LA'=BMRL,
                                'Non-White female'=BFR,
                                'Non-White female LA'=BFRL,
                                'White male first-gen'=WMF,
                                'White male first-gen LA'=WMFL,
                                'White female first-gen'=WFF,
                                'White female first-gen LA'=WFFL,
                                'Non-White male first-gen'=BMF,
                                'Non-White male first-gen LA'=BMFL,
                                'Non-White female first-gen'=BWF,
                                'Non-White female first-gen LA'=BWFL)

students <- glht(intro_m_final, contrast_forms_students)

```


Probabilities (https://sebastiansauer.github.io/convert_logit2prob/)
```{r}
phys_simple_prob <- model_probs(simple2)
phys_simple_prob$seU <- phys_simple_prob$UL-phys_simple_prob$q
phys_simple_prob$seL <- phys_simple_prob$LL-phys_simple_prob$q

phys_courses_prob <- model_probs(courses)
phys_courses_prob$seU <- phys_courses_prob$UL-phys_courses_prob$q
phys_courses_prob$seL <- phys_courses_prob$LL-phys_courses_prob$q
phys_courses_prob <- rbind(phys_simple_prob, phys_courses_prob)
phys_courses_prob$course <- factor(c("All", "All", "204A", "204A", "204B", "204B", "204C", "204C", "202A", "202A", "202B", "202B"))
phys_courses_prob$Instruction <- factor(c("Traditional", "LA","Traditional", "LA", "Traditional", "LA", "Traditional", "LA", "Traditional", "LA", "Traditional", "LA"))
phys_courses_prob$math <- factor(c("Overall", "Overall",  "Calculus-based", "Calculus-based", "Calculus-based", "Calculus-based", "Calculus-based", "Calculus-based", "Algebra-based", "Algebra-based", "Algebra-based", "Algebra-based"))
phys_courses_prob$semester <- factor(c("All sem.", "All sem.", "1st sem.", "1st sem.", "2nd sem.", "2nd sem.", "3rd sem.", "3rd sem.", "1st sem.", "1st sem.", "2nd sem.", "2nd sem."))

overall_prob <- model_probs(overall)
overall_prob$seU <- overall_prob$UL-overall_prob$q
overall_prob$seL <- overall_prob$LL-overall_prob$q
overall_prob$demo <- factor(c("Overall", "Overall"))
overall_prob$demo_fg <- factor(c("Overall", "Overall"))
overall_prob$Race <- factor(c("Overall", "Overall"))
overall_prob$Instruction <- factor(c("Traditional", "LA"))
overall_prob$FG_Status <- factor(c("Overall", "Overall"))

phys_students_prob <- model_probs(students)
phys_students_prob$seU <- phys_students_prob$UL-phys_students_prob$q
phys_students_prob$seL <- phys_students_prob$LL-phys_students_prob$q
phys_students_prob$demo <- factor(c("White Men", "White Men", "White Women", "White Women", "Men of Color", "Men of Color", "Women of Color", "Women of Color", "White Men", "White Men", "White Women", "White Women", "Men of Color", "Men of Color", "Women of Color", "Women of Color"))
phys_students_prob$demo_fg <- factor(c("White Men", "White Men", "White Women", "White Women", "Men of Color", "Men of Color", "Women of Color", "Women of Color", "First Gen. White Men", "First Gen. White Men", "First Gen. White Women", "First Gen. White Women", "First Gen. Men of Color", "First Gen. Men of Color", "First Gen. Women of Color", "First Gen. Women of Color"))
phys_students_prob$Race <- factor(c("White", "White", "White", "White", "POC", "POC", "POC", "POC", "White", "White", "White", "White", "POC", "POC", "POC", "POC"))
phys_students_prob$Instruction <- factor(c("Traditional", "LA", "Traditional", "LA", "Traditional", "LA", "Traditional", "LA", "Traditional", "LA", "Traditional", "LA", "Traditional", "LA", "Traditional", "LA"))
phys_students_prob$FG_Status <- factor(c("Non-FG", "Non-FG", "Non-FG", "Non-FG", "Non-FG", "Non-FG", "Non-FG", "Non-FG", "FG", "FG", "FG", "FG", "FG", "FG", "FG", "FG"))
phys_students_prob2 <- rbind(phys_students_prob,overall_prob)
```

Equality of outcomes prob 
```{r}

eq_out_prob <- data.frame(group=rownames(phys_students_prob),
             q = ifelse(phys_students_prob$Instruction=="LA",phys_students_prob$q-phys_students_prob$q[[2]],
               phys_students_prob$q-phys_students_prob$q[[1]]),
             demo = phys_students_prob$demo,
             Race = phys_students_prob$Race,
             Instruction = phys_students_prob$Instruction,
             FG_Status = phys_students_prob$FG_Status,
             seU = ifelse(phys_students_prob$Instruction=="LA", ((phys_students_prob$seU[[2]])^2+(phys_students_prob$seL)^2)^0.5, ((phys_students_prob$seU[[1]])^2+(phys_students_prob$seL)^2)^0.5),
             seL = ifelse(phys_students_prob$Instruction=="LA", ((phys_students_prob$seL[[2]])^2+(phys_students_prob$seU)^2)^0.5, ((phys_students_prob$seL[[1]])^2+(phys_students_prob$seU)^2)^0.5))
eq_out_prob$UL <- eq_out_prob$q+1.96*eq_out_prob$seU
eq_out_prob$LL <- eq_out_prob$q-1.96*eq_out_prob$seL

             #LL=  phys_students_prob$q-phys_students_prob$q[[1]])
      
#SE = (SE_a^2 + SE_b^2 + 2cov_ab)^0.5 when N_a = N_b
# https://en.wikipedia.org/wiki/Propagation_of_uncertainty

```

Equity of individuality prob
```{r}

eq_ind_prob <-  data.frame(group=rownames(phys_students_prob),
             demo = phys_students_prob$demo,
             Race = phys_students_prob$Race,
             Instruction = ifelse(phys_students_prob$Instruction=="LA",1,0),
             FG_Status = phys_students_prob$FG_Status,
             q= phys_students_prob$q,
             seU= phys_students_prob$seU,
             seL= phys_students_prob$seL)
eq_ind_prob <- eq_ind_prob %>% group_by(demo,FG_Status) %>% mutate(qdiff = q - q[[1]])
eq_ind_prob <- eq_ind_prob %>% group_by(demo,FG_Status) %>% mutate(seUL = ((seL[[1]])^2+(seU)^2)^0.5)
eq_ind_prob <- eq_ind_prob %>% group_by(demo,FG_Status) %>% mutate(seLL = ((seU[[1]])^2+(seL)^2)^0.5)
eq_ind_prob$UL <- eq_ind_prob$qdiff + 1.96*eq_ind_prob$seUL
eq_ind_prob$LL <- eq_ind_prob$qdiff - 1.96*eq_ind_prob$seLL

eq_ind_prob <- eq_ind_prob[which(1:nrow(eq_ind_prob) %% 2 == 0) , ]
```

calculate covariance and errors
```{r}
WMcomp <- summary
est_intro <- summary(intro_m_final)
cov_intro <- vcov(glht(intro_m_final))

pool_and_cov_diffwm <- function(x,y){
  get.est <- foreach(i=1, .combine=rbind) %do% {
  sxp3 <- summary(intro_m_final)
  covp3 <- vcov(glht(intro_m_final))
  data.frame(imp=i, 
             group=rownames(sxp3$linfct),
             d = sxp3$test$coefficients, 
             var.d = (sxp3$test$sigma)^2,
             cov = covp3)
}


p3est <- get.est %>% group_by(group) %>% 
                  summarise(Q = mean(d), 
                            U = mean(var.d), 
                            B = var(d), 
                            T = U + ((1+1/max(imp))*B), 
                            LCL = Q - 1.96*sqrt(T), 
                            UCL = Q + 1.96*sqrt(T),
                            SE = sqrt(T)) 
p3est$race <- word(p3est$group, 1)
p3est$gender <- word(p3est$group, 2)
p3est$Instruction <- word(p3est$group, 3)
p3est$race_gender <- paste(p3est$race,p3est$gender, sep= " ")

return <- p3est}
  
  
  
  
  

get.est <- function(x,y){
  get.est <-data.frame(
             d = est_intro$coefficients, 
             var.d = (est_intro$sigma)^2,
             cov = cov_intro)}

intro_est <- get.est %>% 
                  summarise(Q = mean(d), 
                            U = mean(var.d), 
                            B = var(d), 
                            T = U + ((1+1/max(imp))*B), 
                            LCL = Q - 1.96*sqrt(T), 
                            UCL = Q + 1.96*sqrt(T),
                            SE = sqrt(T),
                            cvwm = mean(cov.White.Male.Lecture)) 
p3est$race <- word(p3est$group, 1)
p3est$gender <- word(p3est$group, 2)
p3est$Instruction <- word(p3est$group, 3)
p3est$race_gender <- paste(p3est$race,p3est$gender, sep= " ")
return <- p3est}

function(x,y){
  get.est <- foreach(i=1:10, .combine=rbind) %do% {
  sxp3 <- summary(glht(x[[i]], linfct=y)) #specifically for post3
  covp3 <- vcov(glht(x[[i]], linfct=y))
  data.frame(imp=i, 
             group=rownames(sxp3$linfct),
             d = sxp3$test$coefficients, 
             var.d = (sxp3$test$sigma)^2,
             cov = covp3)
}

```


Course plot
```{r}
phys_courses_prob$math <- factor(phys_courses_prob$math,levels = c("Overall", "Algebra-based", "Calculus-based"))

courseplot2 <- ggplot(phys_courses_prob, aes(x=semester, y=q, fill=Instruction)) + geom_bar(stat = "identity", position= "dodge") + geom_errorbar(aes( ymax= UL, ymin=LL), position="dodge") +
  theme(legend.position = "bottom", axis.title.x=element_blank(), axis.text.x=element_text(angle=90), axis.ticks.x=element_blank(),panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"),text = element_text(size=16, color = "black")) +
  ylab("DFW probability") + xlab("Semester") + facet_grid(.~math, scales = "free", space = "free")

plot(courseplot2)
ggsave("courses.png", plot= last_plot(), dpi=300, width = 6, height = 4, units = "in", device = "png")
```

Student plot
```{r}
phys_students_prob2$FG_Status <- factor(phys_students_prob2$FG_Status,levels = c("Overall", "Non-FG", "FG"))
phys_students_prob2$FG_Status_Full <- c("Non-First-Generation","Non-First-Generation","Non-First-Generation","Non-First-Generation","Non-First-Generation","Non-First-Generation","Non-First-Generation","Non-First-Generation", "First-Generation","First-Generation","First-Generation","First-Generation","First-Generation","First-Generation","First-Generation","First-Generation", "Overall","Overall")
phys_students_prob2$FG_Status_Full <- factor(phys_students_prob2$FG_Status_Full,levels = c("Overall", "Non-First-Generation", "First-Generation"))
phys_students_prob2$demo_fg <- factor(phys_students_prob2$demo_fg,levels = c("Men of Color", "Women of Color", "White Men","White Women"))

studentplot <- ggplot(phys_students_prob2, aes(x=reorder(demo, new.order=c("Overall", "Men of Color", "Women of Color", "White Men","White Women")),  y=q, fill=Instruction)) + geom_bar(stat = "identity", position= "dodge") + geom_errorbar(aes( ymax= UL, ymin=LL), position="dodge") +
  theme(legend.position = "bottom", axis.title.x=element_blank(), axis.text.x=element_text(angle=90), axis.ticks.x=element_blank(),panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"),text = element_text(size=16, color = "black")) + ylab("Predicted DFW rate") + facet_grid(.~FG_Status_Full, scales = "free", space = "free")

plot(studentplot)
ggsave("students.png", plot= last_plot(), dpi=300, width = 12, height = 6, units = "in", device = "png")
ggsave("students.narrow.png", plot= last_plot(), dpi=300, width = 6, height = 6, units = "in", device = "png")
```

Equality of outcome plot
#```{r}
eq_out_prob$demo_fg <- factor(c("White Men", "White Men", "White Women", "White Women", "Men of Color", "Men of Color", "Women of Color", "Women of Color", " White Men", " White Men", " White Women", " White Women", " Men of Color", " Men of Color", " Women of Color", " Women of Color"))

eq_out_prob <- eq_out_prob[-c(1,2),]

eq_out_plot <- ggplot(eq_out_prob, aes(x=reorder(demo_fg, new.order=c(" Men of Color", " White Men", "Men of Color", " Women of Color", " White Women", "Women of Color", "White Women")), y=q, fill=FG_Status)) + geom_bar(stat = "identity", position= "dodge") + coord_flip()+
  theme(legend.position = "bottom", axis.title.x=element_blank(), axis.text.x=element_text(angle=90), axis.ticks.x=element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"),text = element_text(size=16, color = "black")) + facet_grid(.~Instruction, scales = "free", space = "free") + xlab("") + scale_y_continuous(breaks=c(-0.1,-.05,0,0.05,.1,.15), limits=c(-0.1,0.2))

plot(eq_out_plot)
ggsave("eq_out_plot.png", plot= last_plot(), dpi=300, width = 7, height = 6, units = "in", device = "png")

#```

equity of individuality plot
```{r}
eq_ind_plot <- ggplot(eq_ind_prob, aes(x=reorder(demo, new.order=c("Men of Color", "Women of Color", "White Men","White Women")), y=qdiff, fill=FG_Status)) + geom_bar(stat = "identity", position= "dodge") + 
  theme(legend.position = "bottom", axis.title.x=element_blank(), axis.text.x=element_text(angle=90), axis.ticks.x=element_blank(),panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"),text = element_text(size=16, color = "black")) + ylab("Change in predicted DFW rate")

plot(eq_ind_plot)
ggsave("eq_ind_plot.png", plot= last_plot(), dpi=300, width = 7, height = 6, units = "in", device = "png")

```

#Descriptive statistics
Jayson's code to do the descriptives
```{r}
for_descriptives_df <- red_introPhys[c("DDFW","FG_Status","Female","algebra","nonwhite","LA","race_Non-Resident Alien","race_Unknown","NULLGender")]
descriptives_algebra <- for_descriptives_df %>% group_by(FG_Status, Female, algebra, nonwhite, LA, NULLGender) %>%
  summarise(N = length(DDFW),
            DDFW= sum(DDFW))
descriptives_all <- for_descriptives_df %>% group_by(FG_Status, Female, nonwhite, LA, NULLGender) %>%
  summarise(N = length(DDFW),
            DDFW= sum(DDFW))

descriptives_all_NRA <- for_descriptives_df %>% group_by(FG_Status, Female, nonwhite, LA,`race_Non-Resident Alien`, NULLGender) %>%
  summarise(N = length(DDFW),
            DDFW= sum(DDFW))

red_introPhys$race_Unknown
#write.csv(descriptives_all, file="descriptives_table.csv")
write.csv(descriptives_all_NRA, file="descriptives_table_NRA.csv")

descriptives_all_NRA <- for_descriptives_df %>% group_by(FG_Status, Female, nonwhite, LA,race_Unknown, NULLGender) %>%
  summarise(N = length(DDFW),
            DDFW= sum(DDFW))

#mytable <- xtabs(~`race_Non-Resident Alien`+ `race_American Indian/Alaskan Native` + race_Asian + `race_Black/African American` + `race_Hispanic/Latino` + race_NULL + `race_Pacific Islander` + `race_Two or More Races` + race_White + race_Unknown, data=red_introPhys)
#temp <- ftable(mytable)
#write.csv(temp, file="race sample sizes.csv") # This did not work.

sum(red_introPhys$`race_American Indian/Alaskan Native`)
sum(red_introPhys$race_Asian)
sum(red_introPhys$`race_Black/African American`)
sum(red_introPhys$`race_Hispanic/Latino`)
sum(red_introPhys$`race_Non-Resident Alien`)
sum(red_introPhys$race_NULL)
sum(red_introPhys$`race_Pacific Islander`)
sum(red_introPhys$`race_Two or More Races`)
sum(red_introPhys$race_Unknown)
sum(red_introPhys$race_White)
sum(red_introPhys$race)
sum(red_introPhys$Female=="1" & red_introPhys$Number_202A=="1")
sum(red_introPhys$Female=="1" & red_introPhys$Number_202A=="1")/ sum(red_introPhys$Number_202A=="1")
sum(red_introPhys$gender_Male=="1" & red_introPhys$Number_202A=="1")
sum(red_introPhys$gender_Male=="1" & red_introPhys$Number_202A=="1")/ sum(red_introPhys$Number_202A=="1")
sum(red_introPhys$gender_NULL=="1" & red_introPhys$Number_202A=="1")
sum(red_introPhys$gender_NULL=="1" & red_introPhys$Number_202A=="1")/ sum(red_introPhys$Number_202A=="1")
sum(red_introPhys$Female=="1" & red_introPhys$Number_204A=="1")
sum(red_introPhys$Female=="1" & red_introPhys$Number_204A=="1")/ sum(red_introPhys$Number_204A=="1")
sum(red_introPhys$gender_Male=="1" & red_introPhys$Number_204A=="1")
sum(red_introPhys$gender_Male=="1" & red_introPhys$Number_204A=="1")/ sum(red_introPhys$Number_204A=="1")
sum(red_introPhys$gender_NULL=="1" & red_introPhys$Number_204A=="1")
sum(red_introPhys$gender_NULL=="1" & red_introPhys$Number_204A=="1")/ sum(red_introPhys$Number_204A=="1")
sum(red_introPhys$gender_Male=="1")/count(red_introPhys)
sum(red_introPhys$gender_Female=="1")/count(red_introPhys)
sum(red_introPhys$gender_NULL=="1")/count(red_introPhys)
```


```{r}
desc_gen <- freq(red_phys$gender)
desc_race <- freq(red_phys$race)
desc_FG_Status <- freq(red_phys$FG_Status)
desc_gen
desc_race
desc_FG_Status

desc <- red_phys %>% group_by(Number) %>% summarise(N = length(DDFW),
             sections = length(unique(courID)),
             instructors = length(unique(instructor)),
              DDFW_mean = mean(DDFW),
             DDFW_N = sum(DDFW),
             LA_mean = mean(LA),
             LA_N = sum(LA),
             White_mean = mean(race_White),
             White_N = sum(race_White),
             nonwhite_mean = mean(nonwhite),
             nonwhite_N = sum(nonwhite),
             FG_Status_mean = mean(FG_Status),
             FG_Status_N = sum(FG_Status),
             Male_mean = mean(gender_Male),
             Male_N = sum(gender_Male),
             female_mean = mean(gender_Female),
             female_N = sum(gender_Female))

total <- red_phys %>% summarise(N = length(DDFW),
             sections = length(unique(courID)),
             instructors = length(unique(instructor)),
              DDFW_mean = mean(DDFW),
             DDFW_N = sum(DDFW),
             LA_mean = mean(LA),
             LA_N = sum(LA),
             White_mean = mean(race_White),
             White_N = sum(race_White),
             nonwhite_mean = mean(nonwhite),
             nonwhite_N = sum(nonwhite),
             FG_Status_mean = mean(FG_Status),
             FG_Status_N = sum(FG_Status),
             Male_mean = mean(gender_Male),
             Male_N = sum(gender_Male),
             female_mean = mean(gender_Female),
             female_N = sum(gender_Female))
total$Number <- "total"
total <- total[c(18,1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17)]



desc_intro <-red_introPhys %>% summarise(N = length(DDFW),
                                         sections = length(unique(courID)),
             instructors = length(unique(instructor)),
              DDFW_mean = mean(DDFW),
             DDFW_N = sum(DDFW),
             LA_mean = mean(LA),
             LA_N = sum(LA),
             White_mean = mean(race_White),
             White_N = sum(race_White),
             nonwhite_mean = mean(nonwhite),
             nonwhite_N = sum(nonwhite),
             FG_Status_mean = mean(FG_Status),
             FG_Status_N = sum(FG_Status),
             Male_mean = mean(gender_Male),
             Male_N = sum(gender_Male),
             female_mean = mean(gender_Female),
             female_N = sum(gender_Female))
desc_intro$Number <- "intro"
desc_intro <- desc_intro[c(18,1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17)]

desc_course <- rbind(desc,total,desc_intro)

desc_withincourse <- red_phys %>% group_by(courID) %>% summarise(N = length(DDFW),
              DDFW_mean = mean(DDFW),
             DDFW_N = sum(DDFW),
             LA_mean = mean(LA),
             LA_N = sum(LA),
             White_mean = mean(race_White),
             White_N = sum(race_White),
             nonwhite_mean = mean(nonwhite),
             nonwhite_N = sum(nonwhite),
             FG_Status_mean = mean(FG_Status),
             FG_Status_N = sum(FG_Status),
             Male_mean = mean(gender_Male),
             Male_N = sum(gender_Male),
             female_mean = mean(gender_Female),
             female_N = sum(gender_Female))

desc_WD <- red_phys %>% group_by(nonwhite) %>% summarise(N = count(WD))
```


```{r}
D1 <- lmer(gain~1 + race_URM + gend_URM + stud_pre_cent + coll + (1|course_id),data=MIdata[[1]])
```

Assumption checking 
```{r}
plot(phys_m_final, cex = .6, pch = 1,
     cex.axis = 1/.6, cex.lab = 1/.6,
     cex.main = 1/.6, mar = c(3, 4.5, 0, 1.5))



par(mfrow = c(2, 2))
plot(phys_m_final)
#linearity: Shouldn't see a pattern
print(plot(phys_m_final, xlab="Fitted Value", ylab="Residual Variance"))

#visual homogeneity of variance
red_202_204$Model.F.Res <- residuals(phys_m_final) #extracts the residuals and places them in a new column in our original data table
red_202_204$Abs.Model.F.Res <-abs(red_202_204$Model.F.Res) #creates a new column with the absolute value of the residuals
#print(boxplot(red_202_204$Model.F.Res ~ red_202_204$courID, xlab = "Course", ylab = "Residuals" ))

Levene.Model.F <- lm(Model.F.Res ~ courID, data=red_202_204) #ANOVA of the residuals
anova(Levene.Model.F) #displays the results: want a p>0.05
#visual homogeneity of variance
boxplot(red_202_204$Model.F.Res ~ red_202_204$courID)

#Assumption of Normality or residuals: want points to be near the line
qqmath(phys_m_final)

#library(hglm)
#library(hglm.data)
#plot.hglm(phys_m_final)

#variables are not correlated to the residuals: want a p-value>0.05
cor.test(resid(phys_m_final), MIdata[[1]]$pre_scor) 
cor.test(resid(phys_m_final), MIdata[[1]]$stud_pre_cent)
#quantitative homogeneity of variance
phys_m_final$Model.F.Res<- residuals(phys_m_final) #extracts the residuals and places them in a new column in our original data table
phys_m_final$Abs.Model.F.Res <-abs(phys_m_final$Model.F.Res) #creates a new column with the absolute value of the residuals
#MIdata[[1]]$Model.F.Res2 <- MIdata[[1]]$Abs.Model.F.Res^2 #squares the absolute values of the residuals to provide the more robust estimate
Levene.Model.F <- lm(Model.F.Res ~ course_id, data=MIdata[[1]]) #ANOVA of the residuals
anova(Levene.Model.F) #displays the results: want a p>0.05
#Levene.Model.F2 <- lm(Model.F.Res2 ~ crse_id, data=MIdata[[1]]) #ANOVA of the squared residuals
#anova(Levene.Model.F2) #displays the results
#visual homogeneity of variance
boxplot(MIdata[[1]]$Model.F.Res ~ MIdata[[1]]$course_id)
#boxplot(MIdata[[1]]$Model.F.Res2 ~ MIdata[[1]]$crse_id)
#Assumption of Normality or residuals: want points to be near the line
qqmath(D1)
```


Equity logits #THIS ISN'T HOW TO DO THIS BECAUSE OF LOGITS
```{r}
#Equity of outcomes
#(Int, fg,LA,fem,nwhi, fg:LA,fem:LA,nwhi:LA, fg:nwhi)
WMR =   c(0, 0,0,0,0, 0,0,0, 0) #White, male, returning, no LA
WMRL =  c(0, 0,0,0,0, 0,0,0, 0) #White, male, returning, LA
BMR =   c(0, 0,0,0,1, 0,0,0, 0) 
BMRL =  c(0, 0,0,0,1, 0,0,1, 0)
WFR =   c(0, 0,0,1,0, 0,0,0, 0) 
WFRL =  c(0, 0,0,1,0, 0,1,0, 0)
WMF =   c(0, 1,0,0,0, 0,0,0, 0) 
WMFL =  c(0, 1,0,0,0, 1,0,0, 0)
WFF =   c(0, 1,0,1,0, 0,0,0, 0) 
WFFL =  c(0, 1,0,1,0, 1,1,0, 0)
BFR =   c(0, 0,0,1,1, 0,0,0, 0) 
BFRL =  c(0, 0,0,1,1, 0,1,1, 0)
BMF =   c(0, 1,0,0,1, 0,0,0, 1) 
BMFL =  c(0, 1,0,0,1, 1,0,1, 1)
BWF =   c(0, 1,0,1,1, 0,0,0, 1) 
BWFL =  c(0, 1,0,1,1, 1,1,1, 1)

fit.contrast(phys_m_final,x,BMR)
glh.test(phys_m_final,BMR)

eqout <- rbind (                'White Women'=WFR,
                                'White Women LA'=WFRL,                                
                                'Men of Color'=BMR,
                                'Men of Color LA'=BMRL,
                                'Women of Color'=BFR,
                                'Women of Color LA'=BFRL,
                                'First-gen White Men'=WMF,
                                'First-gen White Men LA'=WMFL,
                                'First-gen White Women'=WFF,
                                'First-gen White Women LA'=WFFL,
                                'First-gen Men of Color'=BMF,
                                'First-gen Men of Color LA'=BMFL,
                                'First-gen Women of Color'=BWF,
                                'First-gen Women of Color LA'=BWFL)

names(eqout) <- c("Int", "fg", "LA", "fem", "POC", "fg:LA", "fem:LA", "POC:LA", "fg:POC")

eq_out <- glht(intro_m_final, eqout)

```

